---
title: "Workplan"
author: "Matt Grainger"
date: "13 2 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mapping to Darwin Core

There have been several attempts to use r to map to and read from Darwin core format (e.g. https://github.com/inbo/dwc-in-R; https://ropensci.org/technotes/2016/12/23/finch-release/). Using these as a basis we can build functions to ensure automation (or reduce the need for coding knowledge as much as possible). 


# Example from https://inbo.github.io/dwc-in-R/dwc-mapping.html 

A raw data file of a checklist of non-native freshwater fish species in Belgium is mapped to Darwin core variables using predominately the tidyverse, magrittr, and janitor packages.  

```{r example from inbo, warning=FALSE, message=FALSE}
# Example from https://inbo.github.io/dwc-in-R/dwc-mapping.html -----------

# Load libraries ----------------------------------------------------------

library(tidyverse) # For data transformation (includes dplyr)
library(magrittr)  # For %<>% pipes
library(readxl)    # For reading excel files
library(janitor)   # For basic cleaning of the data

# Get data ----------------------------------------------------------------

raw_data <- read_excel("Data/AI_2016_Boets_etal_Supplement.xls", 
                                            sheet = "checklist")


# Clean names & add prefix "raw" ------------------------------------------

raw_data %<>% clean_names()    

colnames(raw_data) <- paste0("raw_", colnames(raw_data))


# Keep a copy of the raw data ---------------------------------------------

taxon <- raw_data # Generate taxon by making a copy of raw_data


# Add metadata cols -------------------------------------------------------

taxon %<>% mutate(language = "en")
taxon %<>% mutate(license = "http://creativecommons.org/publicdomain/zero/1.0/")
taxon %<>% mutate(rightsHolder = "Ghent University Aquatic Ecology")
taxon  %<>% mutate(datasetID = "https://doi.org/10.15468/yxcq07")
taxon %<>% mutate(datasetName = "Inventory of alien macroinvertebrates in Flanders, Belgium")

# Add other cols ----------------------------------------------------------
taxon %<>% mutate(kingdom = "Animalia")
taxon %<>% mutate(nomenclaturalCode = "ICZN")


# Unaltered values --------------------------------------------------------

taxon %>% distinct(raw_phylum)
taxon %<>% mutate(taxonID = raw_taxon_id)
taxon %<>% mutate(scientificName = raw_species)
taxon %<>% mutate(family = raw_family)


# Altered values ----------------------------------------------------------

#standard conversion of rawdata

taxon %>% distinct(raw_order)

#Veneroida is a typo. We need to replace this by Venerida:
taxon %<>% 
  mutate(order = recode(raw_order, "Veneroidea" = "Venerida")) 

taxon %>%
  select(raw_order, order) %>%
  group_by_all() %>% 
  summarize()

taxon %>% distinct(raw_phylum)
#However, Crustacea is not a phylum but a subphylum. The phylum to which crustaceans belong is Arthropoda. We can correct this with recode():
taxon %<>% mutate (phylum = recode(raw_phylum, "Crustacea" = "Arthropoda"))

#Other use cases (not applicable to this data):
# distribution %<>% mutate(description = recode(raw_d_n,
#                                                 "Ext.?"     = "Ext.",
#                                                 "Ext./Cas." = "Cas.",
#                                                 "Cas.?"     = "Cas.",
#                                                 "Nat.?"     = "Nat.",
#                                                 .missing = ""))
# 
# distribution %<>% mutate(locality = recode(value,
#                                            "AF"     = "Africa (WGSRPD:2)",
#                                            "AM"     = "pan-American",
#                                            "AS"     = "Asia",
#                                            "AS-Te"  = "temperate Asia (WGSRPD:3)",
#                                            "AS-Tr"  = "tropical Asia (WGSRPD:4)",
#                                            "AUS"    = "Australasia (WGSRPD:5)",
#                                            "Cult."  = "cultivated origin",
#                                            "E"      = "Europe (WGSRPD:1)",
#                                            "Hybr."  = "hybrid origin",
#                                            "NAM"    = "Northern America (WGSRPD:7)",
#                                            "SAM"    = "Southern America (WGSRPD:8)",
#                                            "Trop."  = "Pantropical",
#                                            .default = "",
#                                            .missing = ""
# ))


taxon %>% distinct(raw_species)
# Although most raw_species are in fact species, Dreissena rostriformis bugensis is not a species but a subspecies. We need to make this distinction in the mapping process:
taxon %<>% mutate(taxonRank = case_when(
    raw_species == "Dreissena rostriformis bugensis" ~ "subspecies",
    raw_species != "Dreissena rostriformis bugensis" ~ "species")
  )
# compare our newly mapped data with the original data:
  taxon %>% select(raw_species, taxonRank) %>%
  group_by_all() %>% 
  summarize()

  
  
  # Other use cases (not applicable to our data):
  #   invasion_stage %<>% mutate(description = case_when(
  #     raw_status == "A" ~ "casual",
  #     raw_status == "A*" ~ "casual",
  #     raw_status == "N" ~ "established"
  #   ))
  # 
  # distribution %<>% 
  #   mutate(Flanders = case_when(
  #     raw_presence_fl == "X" & (is.na(raw_presence_br) | raw_presence_br == "?") & (is.na(raw_presence_wa) | raw_presence_wa == "?") ~ "S",
  #     raw_presence_fl == "?" ~ "?",
  #     is.na(raw_presence_fl) ~ "NA",
  #     TRUE ~ "M")) %>%
  #   mutate(Brussels = case_when(
  #     (is.na(raw_presence_fl) | raw_presence_fl == "?") & raw_presence_br == "X" & (is.na(raw_presence_wa) | raw_presence_wa == "?") ~ "S",
  #     raw_presence_br == "?" ~ "?",
  #     is.na(raw_presence_br) ~ "NA",
  #     TRUE ~ "M")) %>%
  #   mutate(Wallonia = case_when(
  #     (is.na(raw_presence_fl) | raw_presence_fl == "?") & (is.na(raw_presence_br) | raw_presence_br == "?") & raw_presence_wa == "X" ~ "S",
  #     raw_presence_wa == "?" ~ "?",
  #     is.na(raw_presence_wa) ~ "NA",
  #     TRUE ~ "M")) %>%
  #   mutate(Belgium = case_when(
  #     raw_presence_fl == "X" | raw_presence_br == "X" | raw_presence_wa == "X" ~ "S", # One is "X"
  #     raw_presence_fl == "?" | raw_presence_br == "?" | raw_presence_wa == "?" ~ "?" # One is "?"
  #   ))  
  
  
  

# Remove rawdata cols -----------------------------------------------------

taxon %<>% select(-starts_with("raw_"))  

head(taxon)  


# save file ---------------------------------------------------------------

#dwc_taxon_file = "../data/processed/taxon.csv"

#write.csv(taxon, file = dwc_taxon_file, na = "", row.names = FALSE, fileEncoding = "UTF-8")

```



<script src="https://hypothes.is/embed.js" async></script>